\chapter{Background}\label{ch:background}

This thesis and its system mostly concerns itself with the topic of DevOps.
What DevOps means and why it is relevant is explained further in this chapter.
More specifically, the subtopics of automation, CI/CD pipelines and testing are looked at.
Concerning the topic of testing, variability-testing is the most relevant form of testing.

\section{DevOps}\label{sec:devops}

The term DevOps often refers to the stages of a product that are between development and release.
Ebert et al.\ describe it as a bridge between "the two worlds of development and operations"\cite{DevOps}.
However, they describe continuous integration and continuous delivery as tools used by DevOps practices rather than being such practices themselves.
They describe it as tools which allow developers to release and deploy a piece of software in a short timeframe.

Leite et al., on the other side, describes DevOps as a set of practices to release software in a short timeframe\cite{ASurveyofDevOpsConceptsandChallenges}.
Furthermore, Chen Lianping describes continuous delivery as an approach rather than a tool\cite{ContinuousDeliveryHugeBenefitsButChallengesToo}.
Apparently, different definitions for the term ''DevOps'' as well as ''Continuous Delivery'' exist.
Therefore, and for the purposes of this paper, the term DevOps is meant as follows.

A set of practices, which do not influence the development of software systems directly.
It does, however, automate large or all parts of the deployment and delivery process.

Since a system to deploy and deliver is needed for a deployment or delivery process, it follows that it is dependent on the development process.
Furthermore, stable and well functioning software is, trivially, more desirable than unstable or malfunctioning software.
It follows that the automation of a delivery process must include ways to ensure, that delivered software is as stable and well functioning as possible.
Otherwise, manual testing and delivery would be needed to ensure this state, which would ultimately defeat the purpose of said process.
In summary, DevOps, for the purposes of this paper, describes the process of automatically testing and delivering a software system.
It must be noticed though, that DevOps practices includes more than just testing and delivering, but these are the focus of this paper.

From this definition two parts become apparent, automatic testing and automatic delivery.
These parts correspond to the practices of continuous integration and continuous delivery.

\section{Continuous integration and continuous delivery}\label{sec:continuous integration and continuous delivery}

Martin Fowler describes ''Continuous integration'' as a software development practice in which code is often committed to a common base repository\cite{ContinuousIntegration}.
''Often'', in this case, meaning multiple times a day supported by automated tests.
Mathias Mayer\cite{ContinuousIntegrationAndItsTools} as well as Shahin et.al.\cite{ContinuousIntegrationDeliveryAndDeploymentASystematicReviewOnApproaches} agree with this definition.
The former one also mentions tools for this practice.
Namely, ''Git'', which is used heavily by this paper, to download, or ''checkout'', the code base to test or deploy.

''Continuous delivery'', on the other hand, is the practice of deploying that code base in fast cycles\cite{ContinuousDeliveryHugeBenefitsButChallengesToo}.
It is valuable to companies, as they can deploy fixes and updates more rapidly.
On the other hand, it creates value for customers, as they do not have to wait a significant amount of time for new features or said fixes.
However, a strong suit of tests is necessary to not introduce even more defects and avoid regression.

\section{Variability testing}\label{sec:variability testing}