\section{Release pipeline}\label{sec:release-pipeline}

In this section, the implementation of the release pipeline is explained.
For each task of the pipeline, the following questions are answered.

\begin{itemize}
    \item What does the task do and how does it do it?
    \item Why does it do it?
    \item What are the configuration options?
    \item What are common error cases and how can they be resolved or mitigated?
\end{itemize}

\subsection{Finding the latest release branch}\label{subsec:finding-the-latest-release-branch}

This task finds the latest release branch using the branches from the GitHub repository of the DTO.
First, the repository is cloned.
Then, the command \\
\verb%git branch -a | grep "${RELEASE_BRANCH_PREFIX}" > ../branches% \\
is used to query all branches, filter the ones with the appropriate prefix and write their names into a file.
The prefix which is used for filtering the names can be configured with the \verb|release_branch_prefix| parameter.
Afterwards a python script, which takes the following parameters, is called.

\begin{table}
    \centering
    \caption{Parameters for the python script finding a release branch}
    \label{tab:py-finding-the-release-branch}
    \begin{tabular}{l|l}
        Parameter & Expected value \\
        \hline
        \verb|--branch-list| & The path to the file containing the filtered branch names \\
        \verb|--release-prefix| & The prefix used to mark release branches. \\
         & It can be configured using the \verb|release_branch_prefix| parameter. \\
        \verb|--csv-prefix| & A prefix for a branch on which the CSV files are generated. \\
        & It can be configured using the \verb|csv_branch_prefix| parameter. \\
        \verb|--helm-prefix| & A prefix for a branch on which the Helm charts are generated. \\
        & It can be configured using the \verb|helm_branch_prefix| parameter. \\
        \verb|--output| & The path to which the results are written. \\
    \end{tabular}
\end{table}

% TODO: format to use verbatim
This python script, using the parameters shown in table \ref{tab:py-finding-the-release-branch}, reads through the list of branches, tries to find a version for each one and then compares the versions found.
It does so, by first removing the strings `{release-prefix}-v` and `{release-prefix}-` from the name.
The resulting name is then split by a dot, i.e., `.`, because it assumes the usage of a semantic versioning scheme.
Since the branches usually contain two version numbers, the major and minor version, the resulting array is extended with `.0` until is length is two.
These steps are necessary to obtain a normalized array for further processing.

The normalized arrays of two branches are then compared to each other.
If an array has a length greater than two, it is not in the expected form for release branches and is discarded.
If an element of the array cannot be parsed to an integer value, it is not in the semver format and is discarded.
If the elements of both arrays can be parsed, they are compared to each other.
If one element is larger then the element with the same index in the other array, it is considered as the later version.
