# release-community-operators

This task releases the CSV bundle for Kubernetes to the community operators repository.
Releasing the CSV files to this repository makes the DTO available on OperatorHub.

It first clones the DTO repository and then creates a fork of the community operators repository.
Forking is done by a python script that can be found at [/scripts/release/fork.py](../../scripts/release/fork.py).
This script takes the following arguments.

| Parameter               | Expected value                                                                                                                                                                         | Default          |
| ----------------------- | -------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- | ---------------- |
| `--token`               | A personal access token for GitHub. It should have the following scopes `admin:org, delete_repo, repo, workflow`. The same token can be used by multiple tasks if these scopes are set |                  |
| `--original-repository` | The original repository which should be forked                                                                                                                                         |                  |
| `--forked-repository`   | The name the forked repository should have                                                                                                                                             |                  |
| `--output-file`         | The path to which the metadata of the fork is written to                                                                                                                               | `fork-data.json` |
| `--retries`             | How often the scripts should retry requests should they fail                                                                                                                           | `10`             |
| `--retry-timeout`       | How long, in seconds, the script waits between request retries.                                                                                                                        | `5`              |

After parsing the arguments, an instance of `GitHubApi`, a class found in [/scripts/release/github/github_api.py](../../scripts/release/github/github_api.py), is created using `{forked-repository}` and `{token}`.
This class implements some REST requests GitHub's API provides.
First, the `exists` function is called, which returns `True` if the forked repository already exists.
To ensure a consistent environment, forked repositories are deleted if they already exist.
Therefore, if the `exist` function returns true, the fork is deleted by calling `delete_repository`, which returns `True` if the request succeeds or `False` if it doesnt.
Additionally, the response from the GitHub API is also returned.

If the forked repository does not exist or has successfully been deleted, it is recreated.
This is done by calling the `fork` function and passing `{original-repository}` as a parameters.
On a success, a check is made, if the forked repository has the expected name `{forked-repository}`.
If it does not, the repository is renamed by instantiating `GitHubApi` with the newly forked repository and calling the `rename` function.
The metadata of the forked repository is then written to `{output-file}`.

In this forked repository, the following folders are created for the new version.
```
${fork_name}/operators/dynatrace-operator/${version}/manifests
${fork_name}/operators/dynatrace-operator/${version}/metadata
```

The Dockerfile located at `{repository}/config/olm/kubernetes/bundle-${version}.Dockerfile` is placed in the version's folder.
The previously mentioned folders, are then populated with the CSV bundles from `{repository}/config/olm/kubernetes/${version}/manifests` and `{repository}/config/olm/kubernetes/${version}/metadata` respectively.
Finally, the changes are added, committed and pushed to the forked repository.

## Purpose

The purpose of this task is to execute all steps necessary to release the CSV bundle to OperatorHub.
This includes forking the repository and following the OperatorHub guidelines.
A pull request or draft pull request is not done to avoid operating on a foreign repository with automatic tasks.

## Configuration options

The parameters can be changed in the [params-dev](../../params-dev.yaml) or [params-prod](../../params-prod.yaml) files.
The following parameters can be used to configure this task.

| Parameter                                 | Effect                                                                                               | Production default                       | Development default                      |
| ----------------------------------------- | ---------------------------------------------------------------------------------------------------- | ---------------------------------------- | ---------------------------------------- |
| `k8s_community_operators_repository`      | Defines the repository that should be forked. I.e., defines the value of `{original-repository}`     | `k8s-operatorhub/community-operators`    | `k8s-operatorhub/community-operators`    |
| `k8s_community_operators_repository_fork` | Defines the name the forked repository should have. I.e., defines the value of `{forked-repository}` | `dt-team-kubernetes/community-operators` | `dt-team-kubernetes/community-operators` |

## Common error cases

### The fork is not created properly

1. Check if the GitHub token is set correctly. It must not be expired and have the necessary scopes.
2. Check if the correct `k8s_community_operators_repository` is targeted.
3. Check if the correct fork, i.e. `k8s_community_operators_repository_fork`, is targeted

### The forked repository could not be found

1. Check that the name of the fork, i.e. `k8s_community_operators_repository_fork`, is set correctly.
2. Sometimes, when a fork is created by the GitHub API, the original name is suffixed by a counter, if the fork already existed on the target account.
   The python script that forks the repository should take care of renaming it, but if it fails to do so, the task may not continue successfully.
