\subsection{Release Rancher}\label{subsec:release-rancher}

This task releases the DTO helm chart to the Rancher repository for partner charts.
It first clones the DTO repository and then create a fork of Rancher's repository.
Forking is done by a python script that can be found at [/scripts/release/fork.py](../../scripts/release/fork.py).
This script takes the following arguments.

| Parameter               | Expected value                                                                                                                                                                         | Default          |
| ----------------------- | -------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- | ---------------- |
| `--token`               | A personal access token for GitHub. It should have the following scopes `admin:org, delete_repo, repo, workflow`. The same token can be used by multiple tasks if these scopes are set |                  |
| `--original-repository` | The original repository which should be forked                                                                                                                                         |                  |
| `--forked-repository`   | The name the forked repository should have                                                                                                                                             |                  |
| `--output-file`         | The path to which the metadata of the fork is written to                                                                                                                               | `fork-data.json` |
| `--retries`             | How often the scripts should retry requests should they fail                                                                                                                           | `10`             |
| `--retry-timeout`       | How long, in seconds, the script waits between request retries.                                                                                                                        | `5`              |

After parsing the arguments, an instance of `GitHubApi`, a class found in [/scripts/release/github/github_api.py](../../scripts/release/github/github_api.py), is created using `{forked-repository}` and `{token}`.
This class implements some REST requests GitHub's API provides.
First, the `exists` function is called, which returns `True` if the forked repository already exists.
To ensure a consistent environment, forked repositories are deleted if they already exist.
Therefore, if the `exist` function returns true, the fork is deleted by calling `delete_repository`, which returns `True` if the request succeeds or `False` if it doesnt.
Additionally, the response from the GitHub API is also returned.

If the forked repository does not exist or has successfully been deleted, it is recreated.
This is done by calling the `fork` function and passing `{original-repository}` as a parameters.
On a success, a check is made, if the forked repository has the expected name `{forked-repository}`.
If it does not, the repository is renamed by instantiating `GitHubApi` with the newly forked repository and calling the `rename` function.
The metadata of the forked repository is then written to `{output-file}`.

When the fork has been created, the task continues to clone it and check out a new branch `dynatrace-operator-v${version}`.
On this branch, a new directory for the version of the new release is created, committed and pushed.
Then, the contents of `{repository}/config/helm/chart/default/`, i.e., the DTO Helm charts, are copied to this new directory.
In the copy of the `Chart.yaml`, the icon property is then set to `file://../logo.png`, to reference the local file.
Finally, the changes are add, committed and pushed to the forked repository.

## Purpose

The purpose of this task is to execute all steps necessary to release the DTO Helm charts to Rancher.
This includes forking their repository and follow their guidelines.
A pull request or draft pull request is not done to avoid operating on a foreign repository with automatic tasks.
